---
title: "Most_deaths"
output: html_document
---

### Observations

-   Most diseases in the dataset are communicable (C), with Tuberculosis leading in total deaths.
-   Pellagra is the only non-communicable (NC) disease, with significantly fewer deaths compared to others.

---
title: "Disease Deaths Comparison"
output: html_document
date: "2025-05-08"
---

## Analysis of Disease Deaths

This report compares total deaths across different diseases, marking them as communicable (C) or non-communicable (NC) using a bar chart.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)

# Print current working directory for debugging
cat("Current working directory:", getwd(), "\n")
```

### Load and Visualize Data

We load the data from the CSV file and create a bar chart to compare deaths by disease.

```{r data_plot, echo=TRUE}
# Attempt to read the CSV file with error handling
tryCatch({
  data <- read.csv("diseases_with_c_nc.csv")
  
  # Create a bar chart
  ggplot(data, aes(x = reorder(Disease, -Total_Deaths), y = Total_Deaths / 1000000, fill = Communicable)) +
    geom_bar(stat = "identity") +
    labs(title = "Comparison of Total Deaths by Disease",
         x = "Disease", y = "Total Deaths (in Millions)",
         fill = "Communicable (C/NC)") +
    scale_y_continuous(breaks = seq(0, max(data$Total_Deaths / 1000000), by = 0.1)) +  # Clear breaks in millions
    scale_fill_manual(values = c("C" = "skyblue", "NC" = "salmon")) +  # Color by communicability
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability
}, error = function(e) {
  cat("Error: Could not read the file 'disease_deaths.csv'. Please ensure the file exists in the working directory (", getwd(), ") or provide the correct file path.\n")
  stop(e)
})
```
